---
title: "Climate Effects on GDP Growth (Kahn et al., 2021, Energy Economics)"
author: "Samuele Centorrino <scentorrino@proton.me>"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
      base_format: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Climate Effects on GDP Growth (Kahn et al., 2021, Energy Economics)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: ref.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r setup, echo=FALSE, cache = FALSE}
library(MASS)
library(plm)
library(knitr)
library(kableExtra)

library(KMNPRY.2021)

options("scipen"=100, "digits"=5)

col.palette <- c('lightgrey','black','grey50')
```

## Methodology in @kahn2021 {-#sec:kahn}

@kahn2021 start from the following ARDL specification (see @kahn2021, p. 4 eq 1)

\begin{equation}
\Delta y_{i,t} = a_i + \sum_{\ell = 1}^p \varphi_\ell \Delta y_{i,t-\ell} + \sum_{\ell = 0}^p \beta_\ell \Delta \tilde{x}_{i,t-\ell}(m) + \varepsilon_{i,t}, 
(\#eq:eq1kahn)
\end{equation}

where $i$ is a country, $t$ is time, $y_{i,s}$ is the logarithm of real GDP for country $i$ and time $s$, $\tilde{x}_{i,s}$ is a vector of regressors which include climate variables such as temperature and precipitations, taken in deviations from their historical averages^[A more detailed technical justification for this choice is given in Appendix A.1 of @kahn2021], $a_i$ is a country fixed effect, and finally $\Delta$ is used to denote first-differences. $\Delta y_{i,t}$ is thus equal to the growth rate of country $i$ at time $t$ which, for simplicity, I denote $r_{i,t}$. Thus,

\begin{equation}
r_{i,t} = a_i + \sum_{\ell = 1}^p \varphi_\ell r_{i,t-\ell} + \sum_{\ell = 0}^p \beta_\ell \Delta \tilde{x}_{i,t-\ell}(m) + \varepsilon_{i,t}. 
(\#eq:eq1kahnr)
\end{equation}

The ARDL model in equation \@ref(eq:eq1kahnr) can be estimated ``as is". However, it is possible that the growth rate, $r_{i,t}$ and $\Delta \tilde{x}_{i,t-\ell}(m)$ are cointegrated (e.g., a long-run equilibrium relationships between these variables may exist), and therefore one can estimate an **error correction version of this model** \citep[see][p. 526]{pesaran2015}. That is,

\begin{equation}
\begin{aligned}
r_{i,t} =& a_i + \sum_{\ell = 1}^p \varphi_\ell r_{i,t-\ell} + \sum_{\ell = 0}^p\Delta \tilde{x}_{i,t-\ell}(m)\beta_\ell  + \varepsilon_{i,t}\\
\Delta r_{i,t}=& a_i - r_{i,t -1} + \sum_{\ell = 1}^p \varphi_\ell r_{i,t-\ell} + \sum_{\ell = 0}^p \Delta \tilde{x}_{i,t-\ell}(m)\beta_\ell  + \varepsilon_{i,t}\\
\Delta r_{i,t}=& a_i - \underbrace{ \left( 1 - \sum_{\ell = 1}^p \varphi_\ell \right)}_{\phi} r_{i,t -1} + \sum_{\ell = 1}^{p-1} \gamma_{\ell} \Delta r_{i,t-\ell} + \sum_{\ell = 0}^p \Delta \tilde{x}_{i,t-\ell}(m)\beta_\ell  + \varepsilon_{i,t}\\
\Delta r_{i,t}=& a_i - \phi \left( r_{i,t -1} - \Delta \tilde{x}_{i,t}(m)\underbrace{\frac{\sum_{\ell = 0}^p\beta_\ell}{\phi}}_{\theta} \right)  + \sum_{\ell = 1}^{p-1} \gamma_{\ell} \Delta r_{i,t-\ell} \\
& \quad - \sum_{\ell = 1}^p \left(\Delta \tilde{x}_{i,t}(m) -  \Delta \tilde{x}_{i,t-\ell}(m)\right) \beta_\ell  + \varepsilon_{i,t}\\
\Delta r_{i,t}=& a_i - \phi \left( r_{i,t -1} - \Delta \tilde{x}_{i,t}(m)\theta \right) + \sum_{\ell = 1}^{p-1} \gamma_{\ell} \Delta r_{i,t-\ell} + \sum_{\ell = 0}^{p-1} \Delta^2 \tilde{x}_{i,t-\ell}(m)\delta_\ell  + \varepsilon_{i,t}, 
\end{aligned}
(\#eq:errcorr)
\end{equation}
where

- $\gamma_\ell = -(\varphi_{\ell+1} + \dots + \varphi_p)$, for $\ell = 1,\dots,p-1$; 
- $\delta_\ell = -(\beta_{\ell+1} + \dots + \beta_p)$,

and $\Delta^2$ denotes the second order difference of $\tilde{x}$. In this formulation of the model, the coefficients of interest are $\phi$ and $\theta$, where $\theta$ captures the long-run effect of the climate variables on the growth rate, and $\phi$ captures the speed of adjustment to this long-run relationship. It is also worth noticing that the fixed effect $a_i$ is still included in the regression, and that therefore this model cannot be directly estimated by Ordinary Least Squares (OLS). 

## Replication of Empirical Results {-#empres}

I first show that using the model in equation \@ref(eq:errcorr), i.e. the error correction version of the model, is numerically equivalent to estimate directly the model in equation \@ref(eq:eq1kahnr), i.e., what I refer to as the *reduced-form* model. 

Results are reported in Tables \@ref(tab:tab1kahn) and \@ref(tab:kahnreduced) for the half-panel jackknife fixed effect estimator (HPJ-FE) of @chudik2018. Notice that the values of the coefficients are numerically equivalent, but there is some discrepancy in the standard errors. This may due to the fact that Stata and R use different methods to compute standard errors.^[An in-depth discussion about this point is beyond the scope of this note. A simple search on the internet reveals the difficulties in replicating Stata standard errors into R] However, the values obtained are not too far apart and do not change the qualitative conclusions of the paper. 

```{r} 
#| label: load_data
#| results: hide
#| message: false
#| warning: false

#### 3 DATABASES VARY ACCORDING TO THE 
#### THE VALUE OF m USED (20,30,40)
## MA 20 
data("data.kahn.ma20",package="KMNPRY.2021")
pdata.kahn.ma20 <- pdata.frame(data.kahn.ma20, index=c("isocode", "year"))  
## MA 30 
data("data.kahn.ma30",package="KMNPRY.2021")
pdata.kahn.ma30 <- pdata.frame(data.kahn.ma30, index=c("isocode", "year")) 
## MA 40 
data("data.kahn.ma40",package="KMNPRY.2021")
pdata.kahn.ma40 <- pdata.frame(data.kahn.ma40, index=c("isocode", "year")) 
```

```{r err_corr_form_1, results = 'hide'}
formula.ardl.tab11 <- as.formula(dgrowth ~ l1growth + dtemp_plus + dtemp_minus + dprecip_plus + dprecip_minus + l1dgrowth +  l2dgrowth + l3dgrowth + d2temp_plus +  l1d2temp_plus + l2d2temp_plus + l3d2temp_plus + d2temp_minus + l1d2temp_minus + l2d2temp_minus + l3d2temp_minus + d2precip_plus + l1d2precip_plus + l2d2precip_plus + l3d2precip_plus + d2precip_minus + l1d2precip_minus + l2d2precip_minus + l3d2precip_minus)

hpj.p.tab11.ma20 <- hpj.fe(formula.ardl.tab11,pdata.kahn.ma20,"individual")
hpj.p.tab11.ma30 <- hpj.fe(formula.ardl.tab11,pdata.kahn.ma30,"individual")
hpj.p.tab11.ma40 <- hpj.fe(formula.ardl.tab11,pdata.kahn.ma40,"individual")

tab1spec1.ma20 <- ec.coef.se.dm(hpj.p.tab11.ma20)
tab1spec1.ma30 <- ec.coef.se.dm(hpj.p.tab11.ma30)
tab1spec1.ma40 <- ec.coef.se.dm(hpj.p.tab11.ma40)
```

```{r err_corr_form_2, results = 'hide'}
formula.ardl.tab12 <- as.formula(dgrowth ~ l1growth + dtemp_plus + dtemp_minus + l1dgrowth +  l2dgrowth + l3dgrowth + d2temp_plus +  l1d2temp_plus + l2d2temp_plus + l3d2temp_plus + d2temp_minus + l1d2temp_minus + l2d2temp_minus + l3d2temp_minus)

hpj.p.tab12.ma20 <- hpj.fe(formula.ardl.tab12,pdata.kahn.ma20,"individual")
hpj.p.tab12.ma30 <- hpj.fe(formula.ardl.tab12,pdata.kahn.ma30,"individual")
hpj.p.tab12.ma40 <- hpj.fe(formula.ardl.tab12,pdata.kahn.ma40,"individual")

tab1spec2.ma20 <- ec.coef.se.dm(hpj.p.tab12.ma20,numreg = 2)
tab1spec2.ma30 <- ec.coef.se.dm(hpj.p.tab12.ma30,numreg = 2)
tab1spec2.ma40 <- ec.coef.se.dm(hpj.p.tab12.ma40,numreg = 2)
```

```{r tab1kahn, tab.cap="HPJ-FE  Estimator with $m = \\lbrace 20,30,40 \\rbrace$ in Table 1 of @kahn2021 (error correction form, standard errors in brackets).", echo =FALSE}
mymat <- cbind(paste0(round(tab1spec1.ma20$lr.coef,3), " (",round(tab1spec1.ma20$lr.coef.se,3),")"),
               paste0(round(tab1spec1.ma30$lr.coef,3),"(",round(tab1spec1.ma30$lr.coef.se,3),")"),
               paste0(round(tab1spec1.ma40$lr.coef,3),"(",round(tab1spec1.ma40$lr.coef.se,3),")"),
               paste0(formatC(c(tab1spec2.ma20$lr.coef[1:2],0,0,tab1spec2.ma20$lr.coef[3]),digits = 3,format = "f",zero.print="~")," (",formatC(c(tab1spec2.ma20$lr.coef.se[1:2],0,0,tab1spec2.ma20$lr.coef.se[3]),digits = 3,format = "f",zero.print="~"),")"),
               paste0(formatC(c(tab1spec2.ma30$lr.coef[1:2],0,0,tab1spec2.ma30$lr.coef[3]),digits = 3,format = "f",zero.print="~")," (",formatC(c(tab1spec2.ma30$lr.coef.se[1:2],0,0,tab1spec2.ma30$lr.coef.se[3]),digits = 3,format = "f",zero.print="~"),")"),
               paste0(formatC(c(tab1spec2.ma40$lr.coef[1:2],0,0,tab1spec2.ma40$lr.coef[3]),digits = 3,format = "f",zero.print="~")," (",formatC(c(tab1spec2.ma40$lr.coef.se[1:2],0,0,tab1spec2.ma40$lr.coef.se[3]),digits = 3,format = "f",zero.print="~"),")"))
rownames(mymat) <- c("$\\hat{\\theta}_{\\Delta \\tilde{T}^{+}_{it}(m)}$","$\\hat{\\theta}_{\\Delta \\tilde{T}^{-}_{it}(m)}$","$\\hat{\\theta}_{\\Delta \\tilde{P}^{+}_{it}(m)}$","$\\hat{\\theta}_{\\Delta \\tilde{P}^{-}_{it}(m)}$","$\\hat{\\phi}$")
colnames(mymat) <- c("MA 20","MA 30","MA 40","MA 20","MA 30","MA 40")

knitr::kable(mymat,digits=3,align = c("l", rep("c", 6)), label = NA) |>
  kableExtra::add_header_above(c("","Specification 1" = 3,"Specification 2" = 3)) |>
     kableExtra::kable_styling(font_size = 10) 
```

```{r reduced_form_1, results = 'hide'}

## ESTIMATE ONLY THE HALF-PANEL JACKNIFE BUT NOT THE FIXED EFFECT
pdata.kahn.ma20 <- createlagvars(pdata.kahn.ma20, ma = 20)
pdata.kahn.ma30 <- createlagvars(pdata.kahn.ma30, ma = 30)
pdata.kahn.ma40 <- createlagvars(pdata.kahn.ma40, ma = 40)

## LOAD DATA
pdata.kahn.tab5.ma20 <- subset(pdata.kahn.ma20,as.numeric(pdata.kahn.ma20$year) >= 61) 
pdata.kahn.tab5.ma30 <- subset(pdata.kahn.ma30,as.numeric(pdata.kahn.ma30$year) >= 61)
pdata.kahn.tab5.ma40 <- subset(pdata.kahn.ma40,as.numeric(pdata.kahn.ma40$year) >= 61) 

# DIRECT SPECIFICATION 1
formula.ardl.tab11dir <- as.formula(growth ~ l1growth + l2growth + l3growth + l4growth + dtemp_plus + l1dtemp_plus + l2dtemp_plus + l3dtemp_plus + l4dtemp_plus + dtemp_minus + l1dtemp_minus + l2dtemp_minus + l3dtemp_minus + l4dtemp_minus + dprecip_plus + l1dprecip_plus + l2dprecip_plus + l3dprecip_plus + l4dprecip_plus + dprecip_minus + l1dprecip_minus + l2dprecip_minus + l3dprecip_minus + l4dprecip_minus)

hpj.p.tab11dir.ma20 <- hpj.fe(formula.ardl.tab11dir,pdata.kahn.ma20,"individual")
hpj.p.tab11dir.ma30 <- hpj.fe(formula.ardl.tab11dir,pdata.kahn.ma30,"individual")
hpj.p.tab11dir.ma40 <- hpj.fe(formula.ardl.tab11dir,pdata.kahn.ma40,"individual")

tab2spec1.ma20 <- sr.coef.se.dm(hpj.p.tab11dir.ma20)
tab2spec1.ma30 <- sr.coef.se.dm(hpj.p.tab11dir.ma30)
tab2spec1.ma40 <- sr.coef.se.dm(hpj.p.tab11dir.ma40)
```

```{r reduced_form_2, results = 'hide'}

formula.ardl.tab12dir <- as.formula(growth ~ l1growth + l2growth + l3growth + l4growth + dtemp_plus + l1dtemp_plus + l2dtemp_plus + l3dtemp_plus + dtemp_minus + l4dtemp_plus + l1dtemp_minus + l2dtemp_minus + l3dtemp_minus + l4dtemp_minus)

hpj.p.tab12dir.ma20 <- hpj.fe(formula.ardl.tab12dir,pdata.kahn.ma20,"individual")
hpj.p.tab12dir.ma30 <- hpj.fe(formula.ardl.tab12dir,pdata.kahn.ma30,"individual")
hpj.p.tab12dir.ma40 <- hpj.fe(formula.ardl.tab12dir,pdata.kahn.ma40,"individual")

tab2spec2.ma20 <- sr.coef.se.dm(hpj.p.tab12dir.ma20,numreg = 2)
tab2spec2.ma30 <- sr.coef.se.dm(hpj.p.tab12dir.ma30,numreg = 2)
tab2spec2.ma40 <- sr.coef.se.dm(hpj.p.tab12dir.ma40,numreg = 2)
```

```{r kahnreduced, tab.cap = "HPJ-FE  Estimator with $m = \\lbrace 20,30,40 \\rbrace$ in Table 1 of @kahn2021 (reduced form, standard errors in brackets).", echo=FALSE}
mymat <- cbind(paste0(round(tab2spec1.ma20$lr.coef,3), " (",round(tab2spec1.ma20$lr.coef.se,3),")"),
               paste0(round(tab2spec1.ma30$lr.coef,3),"(",round(tab2spec1.ma30$lr.coef.se,3),")"),
               paste0(round(tab2spec1.ma30$lr.coef,3),"(",round(tab2spec1.ma40$lr.coef.se,3),")"),
               paste0(formatC(c(tab2spec2.ma20$lr.coef[1:2],0,0,tab2spec2.ma20$lr.coef[3]),digits = 3,format = "f",zero.print="~")," (",formatC(c(tab2spec2.ma20$lr.coef.se[1:2],0,0,tab2spec2.ma20$lr.coef.se[3]),digits = 3,format = "f",zero.print="~"),")"),
               paste0(formatC(c(tab2spec2.ma30$lr.coef[1:2],0,0,tab2spec2.ma30$lr.coef[3]),digits = 3,format = "f",zero.print="~")," (",formatC(c(tab2spec2.ma30$lr.coef.se[1:2],0,0,tab2spec2.ma30$lr.coef.se[3]),digits = 3,format = "f",zero.print="~"),")"),
               paste0(formatC(c(tab2spec2.ma40$lr.coef[1:2],0,0,tab2spec2.ma40$lr.coef[3]),digits = 3,format = "f",zero.print="~")," (",formatC(c(tab2spec2.ma40$lr.coef.se[1:2],0,0,tab2spec2.ma40$lr.coef.se[3]),digits = 3,format = "f",zero.print="~"),")"))
rownames(mymat) <- c("$\\hat{\\theta}_{\\Delta \\tilde{T}^{+}_{it}(m)}$","$\\hat{\\theta}_{\\Delta \\tilde{T}^{-}_{it}(m)}$","$\\hat{\\theta}_{\\Delta \\tilde{P}^{+}_{it}(m)}$","$\\hat{\\theta}_{\\Delta \\tilde{P}^{-}_{it}(m)}$","$\\hat{\\phi}$")
colnames(mymat) <- c("MA 20","MA 30","MA 40","MA 20","MA 30","MA 40")

knitr::kable(mymat,digits=3,align = c("l", rep("c", 6)), label = NA) |>
     kableExtra::add_header_above(c("","Specification 1" = 3,"Specification 2" = 3)) |>
     kableExtra::kable_styling(font_size = 10) 
```


## Replication of Counterfactual Results {-#counterres}

This section reproduces the results of Section 3 in @kahn2021. I do so by estimating a ARDL model as in equation \@ref(eq:eq1kahnr), where the absolute difference between the temperature and its historical mean is the only exogenous regressor. The historical mean is estimated by using windows of size $m = \lbrace 20,30,40 \rbrace$.

Coefficients from this estimation for $m = 30$ are reported in Table \@ref(tab:kahnmainres), which should be compared to Table 5 in @kahn2021. 

```{r}
#| label: create_scenarios

## ADD SCENARIOS FROM KAHN ET AL. (2021)
data("data.kahn.scen",package = "KMNPRY.2021")
data("datappp",package = "KMNPRY.2021")

data.kahn.scen <- merge(datappp,data.kahn.scen,by = "iso")

pdata.kahn.tab5.ma20 <- pdata.frame(merge(pdata.kahn.tab5.ma20,data.kahn.scen,by = "iso"),
                                    index=c("isocode", "year"))
pdata.kahn.tab5.ma30 <- pdata.frame(merge(pdata.kahn.tab5.ma30,data.kahn.scen,by = "iso"),
                                    index=c("isocode", "year"))
pdata.kahn.tab5.ma40 <- pdata.frame(merge(pdata.kahn.tab5.ma40,data.kahn.scen,by = "iso"),
                                    index=c("isocode", "year"))

formula.ardl <- as.formula(growth ~ l1growth + l2growth + l3growth + l4growth +
                            dtempabs + l1dtempabs + l2dtempabs + l3dtempabs + l4dtempabs)

hpj.p.1.ma20 <- hpj.fe(formula.ardl,pdata.kahn.tab5.ma20,"individual")
hpj.p.1.ma30 <- hpj.fe(formula.ardl,pdata.kahn.tab5.ma30,"individual")
hpj.p.1.ma40 <- hpj.fe(formula.ardl,pdata.kahn.tab5.ma40,"individual")
```

```{r kahnmainres, tab.cap="Table 5 in @kahn2021.", echo=FALSE}
mymat <- cbind(hpj.p.1.ma30$coeff,hpj.p.1.ma30$SE)
rownames(mymat) <- c("$\\hat{\\varphi}_1$","$\\hat{\\varphi}_2$","$\\hat{\\varphi}_3$","$\\hat{\\varphi}_4$","$\\hat{\\beta}_0$","$\\hat{\\beta}_1$","$\\hat{\\beta}_2$","$\\hat{\\beta}_3$","$\\hat{\\beta}_4$")
colnames(mymat) <- c("Coefficient","Std. Err.")

knitr::kable(t(mymat),digits=4,align = c("l", rep("c", 9)), label = NA) |>
     kableExtra::kable_styling(font_size = 10) 
```

I further proceed to obtain the expected losses in GPD by country for a given scenario about the expected evolution of temperatures around their historical mean. 

To do so, I first rewrite equation \@ref(eq:eq1kahnr) in the following form

\begin{equation}
\varphi(L) r_{it} = a_i + \beta(L) \Delta x_{it}(m) + \varepsilon_{it}, 
(\#eq:eq1kahnr2)
\end{equation}

where $\Delta x_{it}(m) = \vert T_{it} - T^\ast_{it-1}(m)\vert$,
\begin{align*}
\varphi(L) =& 1 - \sum_{\ell = 1}^4 \varphi_\ell L^\ell \\
\beta(L)=& \sum_{\ell = 0}^4 \beta_\ell L^\ell,
\end{align*}
and $L$ is the lag operator. That is,
\[
r_{it} = \varphi(1)^{-1} a_i + \psi(L) \Delta x_{it}(m) + \vartheta(L)\varepsilon_{it},
\]
where $\vartheta(L) = \sum_{\ell=0}^\infty \vartheta_\ell L^\ell$, and
\[
\psi(L) = \varphi(L)^{-1} \beta(L) = \sum_{\ell=0}^\infty \psi_\ell L^\ell.
\]
Obtaining an estimator of $\psi$ allows us to forecast the expected value of GDP growth $h$ periods ahead, for a given sequence of observed temperatures (in absolute deviations from their historical mean) between $T$ and $T+h$. 

```{r eval=T,echo=F}
#| label: fs_counterfactuals
n.years      <- 101

## Obtain weights for counterfactual scenarios
##### MA 20
psihat.ma20 <- psicoef(hpj.p.1.ma20)

##### MA 30
psihat.ma30 <- psicoef(hpj.p.1.ma30)

### MA 40
psihat.ma40 <- psicoef(hpj.p.1.ma40)

## Create new database with numeric year
## It does not matter which database you take to generate the new data, as beta and sigma are generated from the raw series of temperatures
newclimatedata <- pdata.kahn.tab5.ma20
newclimatedata$yearmod <- as.numeric(newclimatedata$year) + 1959
newclimatedata$isocode <- as.numeric(newclimatedata$isocode)

## Obtain slope for the trend and variance of the residuals
newclimatedata$beta.Ti  <- numeric(dim(newclimatedata)[1])
newclimatedata$sigma.Ti <- numeric(dim(newclimatedata)[1])

for(i in 1:174){

  tmpmod <- lm(temp ~ yearmod, data = newclimatedata[newclimatedata$isocode == i,])
  newclimatedata$beta.Ti[newclimatedata$isocode == i] <- tmpmod$coefficients[2]
  newclimatedata$sigma.Ti[newclimatedata$isocode == i] <- sqrt(sum(tmpmod$residuals^2)/(length(tmpmod$residuals)-2))
 
}

climate.data2014 <- subset(newclimatedata,newclimatedata$year == 2014)

## Generate weights for cold, hot, poor, and rich countries
# COLD
with(climate.data2014,pppgdpwc <- (1 - hot)* pppgdpw/sum((1 - hot)* pppgdpw))

# HOT
with(climate.data2014,pppgdpwh <- hot* pppgdpw/sum(hot* pppgdpw))

# POOR
with(climate.data2014,pppgdpwp <- poor*pppgdpw/sum(poor*pppgdpw))

# RICH
with(climate.data2014,pppgdpwr <- (1 - poor)*pppgdpw/sum((1 - poor)*pppgdpw))
```

In @kahn2021, the authors consider that temperature follows a model with a linear trend plus noise. Under an additional assumption that the noise is normally distributed, they derive the expected change in temperature between time $T$ and time $T + h$ in closed form, and use it to predict the expected loss in GDP for several scenarios. In particular, they have

\begin{equation}
\begin{aligned}
T_{i, T + j} =& a_{Ti} + b_{Ti,j} (T + j) + v_{Ti,T+j}, \text{ for } j = 1,2,\dots \\
v_{Ti,t} \sim & N\left( 0, \sigma^2_{Ti}\right)\\
E  \vert T_{i, T + j} - T^\ast_{i, T+ j-1}(m)\vert =& \mu_{Ti,j}\left[ \Phi \left( \frac{\mu_{Ti,j}}{\omega_{Ti}}\right) - 
\Phi\left( -\frac{\mu_{Ti,j}}{\omega_{Ti}}\right)\right] + 2 \omega_{Ti}\phi\left( \frac{\mu_{Ti,j}}{\omega_{Ti}}\right),
\end{aligned}
\end{equation}

where $\mu_{Ti,j} = b_{Ti,j} (m+1)/2$, $\omega^2_{Ti}= \sigma^2_{Ti} (1 + 1/m)$, and $\{\phi,\Phi\}$ are the pdf and cdf of a standard normal distribution. 

Let $g_{Ti}(m,b_{Ti,j},\sigma_{Ti}) = E  \vert T_{i, T + j} - T^\ast_{i, T+ j-1}(m)\vert$, the change in the log real GDP per-capita of country $i$ over the horizon $h$, is given by

\begin{equation}
\Delta_{h}(d_i,m) = \sum_{j = 1}^h \psi_{h-j} \left( g_{Ti}(m,b^0_{Ti} + jd_i,\sigma^0_{Ti})  - g_{Ti}(m,b^0_{Ti},\sigma^0_{Ti})\right),
\end{equation}

where $b^0_{Ti}$ and $\sigma^0_{Ti}$ are the trend and the standard deviation of the error term estimated using observations from years 1960-2014.

@kahn2021 consider the uncertainty coming from the estimation of $\psi$, and provide confidence intervals for these parameters. However, they do not provide forecast intervals for the expected loss in GDP. $\hat{\psi}$ is a nonlinear transformation of $\lbrace \hat{\varphi},\hat{\beta} \rbrace$, so that *standard* asymptotic confidence intervals based on asymptotic normality of the estimator are hard to derive in closed form. In order to obtain confidence intervals, I bootstrap the panel with respect to its cross sectional dimension using a dynamic wild bootstrap to allow for serial dependence between countries (@shao2010, and @gao2022). The first $20$ values of $\hat{\psi}$ with their $95\%$ confidence intervals are given in Figure \@ref(fig:fig5kahn), which mimics very closely Figure 5 in the original paper. 

I present here an example with $m = 30$ about how to invoke the function to obtain $95\%$ bootstrap confidence intervals with $B = 499$ replications.

```{r}
#| label: bootstrap_coef

## SET PARAMETERS FOR BOOTSTRAP
set.seed(123)
B <- 499

psihat.b.ma30 <- psicoef.b(pdata.kahn.tab5.ma30, formula.ardl, 
                           hpj.p.1.ma30$coeff,hpj.p.1.ma30$resid, B = B)

psihat025.ma30 <- apply(psihat.b.ma30$psihat,1,function(x) quantile(x,0.025))
psihat975.ma30 <- apply(psihat.b.ma30$psihat,1,function(x) quantile(x,0.975))
``` 


```{r fig5kahn, fig.cap="$\\psi_j$ for $j = 0,1,2,\\dots,20$, and their bootstrapped $95\\%$ confidence intervals.", fig.dim = c(6, 5)}
barCenters <- barplot(psihat.ma30[1:21],xlim = c(0,21),ylim = rev(c(-0.02,0.001)),col = "blue4",cex.axis = 0.5,legend = expression(psi[i]))
segments(barCenters, psihat025.ma30[1:21], barCenters,
            psihat975.ma30[1:21], lwd = 2,col = "red4" )
arrows(barCenters, psihat025.ma30[1:21], barCenters,
            psihat975.ma30[1:21], lwd = 2, angle = 90,
            code = 3, length = 0.1,col = "red4")
``` 

In Table \@ref(tab:tab6kahn), I report the expected loss of GDP growth (in percentage) for the two scenarios considered by @kahn2021. The $95\%$ confidence intervals are constructed via the bootstrapped values of $\hat{\psi}$.

```{r tab6kahn, tab.cap= "Table 6 in @kahn2021 with $m = 30$.", echo = FALSE}
## REPRODUCE THE RESULTS FOR m = 30
m  <- 30
counter26 <- with(climate.data2014,cccounter(m,beta.Ti,sigma.Ti,dim(climate.data2014)[1],psihat.ma30,di_rcp26))
Delta_30_26 <- counter26$Delta
counter85 <- with(climate.data2014,cccounter(m,beta.Ti,sigma.Ti,dim(climate.data2014)[1],psihat.ma30,di_rcp85))
Delta_30_85 <- counter85$Delta

## CONFIDENCE INTERVALS FOR SCENARIOS
Delta_30_26_025 <- matrix(0,dim(climate.data2014)[1],length(psihat.ma30)-1)
Delta_30_85_025 <- matrix(0,dim(climate.data2014)[1],length(psihat.ma30)-1)
Delta_30_26_975 <- matrix(0,dim(climate.data2014)[1],length(psihat.ma30)-1)
Delta_30_85_975 <- matrix(0,dim(climate.data2014)[1],length(psihat.ma30)-1)
for(i in 1:174){
    tmpDelta26 <- matrix(0,B,length(psihat.ma30)-1)
    tmpDelta85 <- matrix(0,B,length(psihat.ma30)-1)
    for(jj in 1:B){
        counter26.b <- with(climate.data2014,cccounter(m,beta.Ti[i],sigma.Ti[i],1,psihat.b.ma30$psihat[,jj],di_rcp26[i]))
        counter85.b <- with(climate.data2014,cccounter(m,beta.Ti[i],sigma.Ti[i],1,psihat.b.ma30$psihat[,jj],di_rcp85[i]))
        tmpDelta26[jj,] <- counter26.b$Delta
        tmpDelta85[jj,] <- counter85.b$Delta
    } 
    Delta_30_26_025[i,] <- apply(tmpDelta26,2,function(x) quantile(x,0.025))
    Delta_30_85_025[i,] <- apply(tmpDelta85,2,function(x) quantile(x,0.025))
    Delta_30_26_975[i,] <- apply(tmpDelta26,2,function(x) quantile(x,0.975))
    Delta_30_85_975[i,] <- apply(tmpDelta85,2,function(x) quantile(x,0.975))
}

## WORLD AVERAGES
wavg_Delta_30_26  <- numeric(length(psihat.ma30)-1)
wavg_Delta_30_85  <- numeric(length(psihat.ma30)-1)
wcavg_Delta_30_26 <- numeric(length(psihat.ma30)-1)
wcavg_Delta_30_85 <- numeric(length(psihat.ma30)-1)
whavg_Delta_30_26 <- numeric(length(psihat.ma30)-1)
whavg_Delta_30_85 <- numeric(length(psihat.ma30)-1)
wravg_Delta_30_26 <- numeric(length(psihat.ma30)-1)
wravg_Delta_30_85 <- numeric(length(psihat.ma30)-1)
wpavg_Delta_30_26 <- numeric(length(psihat.ma30)-1)
wpavg_Delta_30_85 <- numeric(length(psihat.ma30)-1)

for(j in 1:100){

  wavg_Delta_30_26[j] <- sum(climate.data2014$pppgdpw * Delta_30_26[,j])
  wavg_Delta_30_85[j] <- sum(climate.data2014$pppgdpw * Delta_30_85[,j])
  
  wcavg_Delta_30_26[j]<- sum(climate.data2014$pppgdpwc * Delta_30_26[,j])
  wcavg_Delta_30_85[j]<- sum(climate.data2014$pppgdpwc * Delta_30_85[,j])
  
  whavg_Delta_30_26[j]<- sum(climate.data2014$pppgdpwh * Delta_30_26[,j])
  whavg_Delta_30_85[j]<- sum(climate.data2014$pppgdpwh * Delta_30_85[,j])
  
  wravg_Delta_30_26[j]<- sum(climate.data2014$pppgdpwr * Delta_30_26[,j])
  wravg_Delta_30_85[j]<- sum(climate.data2014$pppgdpwr * Delta_30_85[,j])
  
  wpavg_Delta_30_26[j]<- sum(climate.data2014$pppgdpwp * Delta_30_26[,j])
  wpavg_Delta_30_85[j]<- sum(climate.data2014$pppgdpwp * Delta_30_85[,j])
  
}

EUlist <- c('Austria','Belgium','Bulgaria','Croatia','Cyprus','Czech Republic','Denmark','Estonia','Finland','France','Germany','Greece','Hungary','Ireland','Italy','Latvia','Lithuania','Luxembourg','Malta','Netherlands','Poland','Portugal','Romania','Slovakia','Slovenia','Spain','Sweden')
hvec <- c(16,36,86)
worldtab <- rbind(formatC(wavg_Delta_30_26[hvec]*(-100),digits = 2,format = "f"),formatC(wavg_Delta_30_85[hvec]*(-100),digits = 2,format = "f"))

#### CHINA
chinatab <- rbind(formatC(Delta_30_26[climate.data2014$cntry_name %in% "China",hvec]*(-100),digits = 2,format = "f"),formatC(Delta_30_85[climate.data2014$cntry_name == "China",hvec]*(-100),digits = 2,format = "f"))

cilb26 <- apply(rbind(Delta_30_26_975[climate.data2014$cntry_name == "China",hvec],Delta_30_26_025[climate.data2014$cntry_name == "China",hvec]*(-100)),2,min)
ciub26 <- apply(rbind(Delta_30_26_975[climate.data2014$cntry_name == "China",hvec],Delta_30_26_025[climate.data2014$cntry_name == "China",hvec]*(-100)),2,max)
cilb85 <- apply(rbind(Delta_30_85_975[climate.data2014$cntry_name == "China",hvec]*(-100),Delta_30_85_025[climate.data2014$cntry_name == "China",hvec]*(-100)),2,min)
ciub85 <- apply(rbind(Delta_30_85_975[climate.data2014$cntry_name == "China",hvec]*(-100),Delta_30_85_025[climate.data2014$cntry_name == "China",hvec]*(-100)),2,max)
chinatabci <- rbind(paste0("[",formatC(cilb26,digits = 2,format = "f"),",",formatC(ciub26,digits = 2,format = "f"),"]"),paste0("[",formatC(cilb85,digits = 2,format = "f"),",",formatC(ciub85,digits = 2,format = "f"),"]"))
chinatab <- rbind(chinatab[1,],chinatabci[1,],chinatab[2,],chinatabci[2,])

#### EU
eutab <- rbind(formatC(apply(Delta_30_26[climate.data2014$cntry_name %in% EUlist,hvec]*(-100),2,mean),digits = 2,format = "f"),
               formatC(apply(Delta_30_85[climate.data2014$cntry_name %in% EUlist,hvec]*(-100),2,mean),digits = 2,format = "f"))

cilb26 <- apply(rbind(apply(Delta_30_26_975[climate.data2014$cntry_name %in% EUlist,hvec]*(-100),2,mean),apply(Delta_30_26_025[climate.data2014$cntry_name %in% EUlist,hvec]*(-100),2,mean)),2,min)
ciub26 <- apply(rbind(apply(Delta_30_26_975[climate.data2014$cntry_name %in% EUlist,hvec]*(-100),2,mean),apply(Delta_30_26_025[climate.data2014$cntry_name %in% EUlist,hvec]*(-100),2,mean)),2,max)
cilb85 <- apply(rbind(apply(Delta_30_85_975[climate.data2014$cntry_name %in% EUlist,hvec]*(-100),2,mean),apply(Delta_30_85_025[climate.data2014$cntry_name %in% EUlist,hvec]*(-100),2,mean)),2,min)
ciub85 <- apply(rbind(apply(Delta_30_85_975[climate.data2014$cntry_name %in% EUlist,hvec]*(-100),2,mean),apply(Delta_30_85_025[climate.data2014$cntry_name %in% EUlist,hvec]*(-100),2,mean)),2,max)

eutabci <- rbind(paste0("[",formatC(cilb26,digits = 2,format = "f"),",",formatC(ciub26,digits = 2,format = "f"),"]"),paste0("[",formatC(cilb85,digits = 2,format = "f"),",",formatC(ciub85,digits = 2,format = "f"),"]"))

eutab <- rbind(eutab[1,],eutabci[1,],eutab[2,],eutabci[2,])

#### INDIA
cilb26 <- apply(rbind(Delta_30_26_975[climate.data2014$cntry_name == "India",hvec]*(-100),Delta_30_26_025[climate.data2014$cntry_name == "India",hvec]*(-100)),2,min)
ciub26 <- apply(rbind(Delta_30_26_975[climate.data2014$cntry_name == "India",hvec]*(-100),Delta_30_26_025[climate.data2014$cntry_name == "India",hvec]*(-100)),2,max)
cilb85 <- apply(rbind(Delta_30_85_975[climate.data2014$cntry_name == "India",hvec]*(-100),Delta_30_85_025[climate.data2014$cntry_name == "India",hvec]*(-100)),2,min)
ciub85 <- apply(rbind(Delta_30_85_975[climate.data2014$cntry_name == "India",hvec]*(-100),Delta_30_85_025[climate.data2014$cntry_name == "India",hvec]*(-100)),2,max)
indiatab <- rbind(formatC(Delta_30_26[climate.data2014$cntry_name == "India",hvec]*(-100),digits = 2,format = "f"),formatC(Delta_30_85[climate.data2014$cntry_name == "India",hvec]*(-100),digits = 2,format = "f"))
indiatabci <- rbind(paste0("[",formatC(cilb26,digits = 2,format = "f"),",",formatC(ciub26,digits = 2,format = "f"),"]"),paste0("[",formatC(cilb85,digits = 2,format = "f"),",",formatC(ciub85,digits = 2,format = "f"),"]"))
indiatab <- rbind(indiatab[1,],indiatabci[1,],indiatab[2,],indiatabci[2,])

#### RUSSIA
cilb26 <- apply(rbind(Delta_30_26_975[climate.data2014$cntry_name == "Russian Federation",hvec]*(-100),Delta_30_26_025[climate.data2014$cntry_name == "Russian Federation",hvec]*(-100)),2,min)
ciub26 <- apply(rbind(Delta_30_26_975[climate.data2014$cntry_name == "Russian Federation",hvec]*(-100),Delta_30_26_025[climate.data2014$cntry_name == "Russian Federation",hvec]*(-100)),2,max)
cilb85 <- apply(rbind(Delta_30_85_975[climate.data2014$cntry_name == "Russian Federation",hvec]*(-100),Delta_30_85_025[climate.data2014$cntry_name == "Russian Federation",hvec]*(-100)),2,min)
ciub85 <- apply(rbind(Delta_30_85_975[climate.data2014$cntry_name == "Russian Federation",hvec]*(-100),Delta_30_85_025[climate.data2014$cntry_name == "Russian Federation",hvec]*(-100)),2,max)

russiatab <- rbind(formatC(Delta_30_26[climate.data2014$cntry_name == "Russian Federation",hvec]*(-100),digits = 2,format = "f"),formatC(Delta_30_85[climate.data2014$cntry_name == "Russian Federation",hvec]*(-100),digits = 2,format = "f"))
russiatabci <- rbind(paste0("[",formatC(cilb26,digits = 2,format = "f"),",",formatC(ciub26,digits = 2,format = "f"),"]"),paste0("[",formatC(cilb85,digits = 2,format = "f"),",",formatC(ciub85,digits = 2,format = "f"),"]"))
russiatab <- rbind(russiatab[1,],russiatabci[1,],russiatab[2,],russiatabci[2,])

#### UNITED STATES
cilb26 <- apply(rbind(Delta_30_26_975[climate.data2014$cntry_name == "United States",hvec]*(-100),Delta_30_26_025[climate.data2014$cntry_name == "United States",hvec]*(-100)),2,min)
ciub26 <- apply(rbind(Delta_30_26_975[climate.data2014$cntry_name == "United States",hvec]*(-100),Delta_30_26_025[climate.data2014$cntry_name == "United States",hvec]*(-100)),2,max)
cilb85 <- apply(rbind(Delta_30_85_975[climate.data2014$cntry_name == "United States",hvec]*(-100),Delta_30_85_025[climate.data2014$cntry_name == "United States",hvec]*(-100)),2,min)
ciub85 <- apply(rbind(Delta_30_85_975[climate.data2014$cntry_name == "United States",hvec]*(-100),Delta_30_85_025[climate.data2014$cntry_name == "United States",hvec]*(-100)),2,max)

ustab <- rbind(formatC(Delta_30_26[climate.data2014$cntry_name == "United States",hvec]*(-100),digits = 2,format = "f"),formatC(Delta_30_85[climate.data2014$cntry_name == "United States",hvec]*(-100),digits = 2,format = "f"))
ustabci <- rbind(paste0("[",formatC(cilb26,digits = 2,format = "f"),",",formatC(ciub26,digits = 2,format = "f"),"]"),paste0("[",formatC(cilb85,digits = 2,format = "f"),",",formatC(ciub85,digits = 2,format = "f"),"]"))
ustab <- rbind(ustab[1,],ustabci[1,],ustab[2,],ustabci[2,])


mytab <- rbind(c("World",rep("",3)),
               cbind(c("RCP 2.6","RCP 8.5"),formatC(worldtab,digits = 2,format = "f")),
               c("China",rep("",3)),
               cbind(c("RCP 2.6","","RCP 8.5",""),chinatab),
               c("European Union",rep("",3)),
               cbind(c("RCP 2.6","","RCP 8.5",""),eutab),
               c("India",rep("",3)),
               cbind(c("RCP 2.6","","RCP 8.5",""),indiatab),
               c("Russia",rep("",3)),
               cbind(c("RCP 2.6","","RCP 8.5",""),russiatab),
               c("United States",rep("",3)),
               cbind(c("RCP 2.6","","RCP 8.5",""),ustab))
               
colnames(mytab) <- c("~","2030","2050","2100")

knitr::kable(mytab,digits=3,align = c("l", rep("c", 9)),label = NA) |>
    kableExtra::row_spec(c(1,4,9,14,19,24),bold=T,hline_after = T) |>
     kableExtra::kable_styling(font_size = 10) 
```


## Heterogenous effects {-#heteffts}

I finally replicate the results in Table 4 of @kahn2021, where they consider estimation of the model in \@ref(eq:eq1kahnr) with heterogenous coefficients and common correlated effects. In particular, they estimate the model

\begin{equation}
r_{i,t} = a_i + \varphi_{1,i} r_{i,t-\ell} + \sum_{\ell = 0}^p \beta_{\ell,i} \Delta \tilde{x}_{i,t-\ell}(m) + \omega_i r_{w,t-1} + \varepsilon_{i,t},
(\#eq:eq1kahnrhet)
\end{equation}
with $i = 1,\dots,n$. In equation \@ref(eq:eq1kahnrhet), $p = 4$, and I only take one lag of GDP growth as explanatory variable to reduce the number of parameters; $r_{w,t-1}$ denotes the world GDP growth, and controls for common correlated effects among all countries. All coefficients are taken to be heterogeneous across countries. This model is estimated by OLS country-by-country. The mean-group estimator of the long-run effect of temperature on GDP growth is defined as

\begin{equation}
\hat{\theta}_{MG} = \frac{\sum_{\ell = 1}^p \frac{1}{n} \sum_{i = 1}^n \hat{\beta}_{\ell,i}}{1 - \frac{1}{n}\sum_{i = 1}^n \hat{\varphi}_{1,i}}.
\end{equation}

It is known that this estimator can be biased for small values of T, so the authors follow @chudik2018, and construct a half-panel jackknife bias corrected estimator of $\hat{\theta}_{MG}$ as 

\begin{equation}
\hat{\theta}_{HP,MG} = 2\hat{\theta}_{MG} - 0.5 \left( \hat{\theta}_{1, MG} + \hat{\theta}_{2,MG}\right),
(\#eq:mgjkhp)
\end{equation}

where $\hat{\theta}_{1, MG}$ and $\hat{\theta}_{2, MG}$ are the estimators constructed by splitting the panel over the time dimension. The variance of the mean-group half-panel jackknife estimator in \@ref(eq:mgjkhp) is then given by

\begin{equation}
\begin{aligned}
var \left( \hat{\theta}_{HP,MG} \right) =& 4 var \left( \hat{\theta}_{MG} \right)  + 0.25 \left( var \left( \hat{\theta}_{1,MG}\right) + var \left(\hat{\theta}_{2,MG}\right) + cov\left( \hat{\theta}_{1,MG},\hat{\theta}_{2,MG}\right) + cov\left( \hat{\theta}_{2,MG},\hat{\theta}_{1,MG}\right)\right)\\
& \quad - cov\left(\hat{\theta}_{MG},\hat{\theta}_{1,MG}\right) - cov\left(\hat{\theta}_{MG},\hat{\theta}_{2,MG}\right) -cov\left(\hat{\theta}_{1,MG},\hat{\theta}_{MG}\right) - cov\left(\hat{\theta}_{2,MG},\hat{\theta}_{MG}\right).
\end{aligned}
(\#eq:varmgjkhp)
\end{equation}

I restrict the sample to countries with more than 30 non-missing observations for GDP growth, which results in a sample of 130 countries. As in the original paper, I estimate the model in \@ref(eq:eq1kahnrhet) both when $\omega_i$ is restricted to be equal to $0$ and when left unrestricted (a model without and with common correlated effects, respectively).

The function `hpj.het` returns the country-by-country half-panel jackknife estimators of the short-run coefficients for all countries. In the next chunk of code, I provide an example of the implementation of this function for $MA = 30$. The results for other adaptation windows can be reproduced in a similar way.

```{r, label = het.coeff, results = 'hide'}
library(dplyr)

pdata.kahn.tab4.ma30 <- pdata.kahn.tab5.ma30 %>% dplyr::filter(!is.na(growth)) %>%
                                      group_by(isocode) %>%
                                        dplyr::filter(sum(!is.na(growth)) > 30) %>% 
                                            ungroup() %>%
                                              dplyr::filter(!is.na(l1growth)) %>% 
                                              plm::pdata.frame(index=c("iso", "year"))

## Heterogenous estimation without CCE
formula.ardl.tab4.nocce <- as.formula(growth ~ l1growth + dtempabs + l1dtempabs + l2dtempabs + l3dtempabs + l4dtempabs)

het.coeff.nocce <- hpj.het(formula.ardl.tab4.nocce, pdata.kahn.tab4.ma30)

## Heterogenous estimation with CCE
formula.ardl.tab4.cce <- as.formula(growth ~ l1growth + dtempabs + l1dtempabs + l2dtempabs + l3dtempabs + l4dtempabs + l1gg)

het.coeff.cce <- hpj.het(formula.ardl.tab4.cce, pdata.kahn.tab4.ma30[!is.na(pdata.kahn.tab4.ma30$l1gg),])
```

Then the function `lr_het_coef` can be used to compute the long-run bias corrected estimate from equation \@ref(eq:mgjkhp), and its standard error from equation \@ref(eq:varmgjkhp).

Table \@ref(tab:lrcoefbygroup) reports the results of this exercise. The results correspond to those in the original paper with two exceptions. Standard errors differ from those reported in the paper, and I did not manage to replicate the effects for cold and hot countries.

```{r lrcoefbygroup, tab.cap = "Table 4 in @kahn2021 with $m = 30$.", echo = FALSE}
data("country.vars",package="KMNPRY.2021")

all.ma30.nocce <- lr_het_coef(het.coeff.nocce)
cold.ma30.nocce <- lr_het_coef(het.coeff.nocce,subgroup = country.vars$iso[country.vars$cold == 1])
hot.ma30.nocce <- lr_het_coef(het.coeff.nocce,subgroup = country.vars$iso[country.vars$cold == 0])
poor.ma30.nocce <- lr_het_coef(het.coeff.nocce,subgroup = country.vars$iso[country.vars$poor == 1])
rich.ma30.nocce <- lr_het_coef(het.coeff.nocce,subgroup = country.vars$iso[country.vars$aeg20em == 1])

all.ma30.cce <- lr_het_coef(het.coeff.cce)
cold.ma30.cce <- lr_het_coef(het.coeff.cce,country.vars$iso[country.vars$cold == 1])
hot.ma30.cce <- lr_het_coef(het.coeff.cce,country.vars$iso[country.vars$cold == 0])
poor.ma30.cce <- lr_het_coef(het.coeff.cce,country.vars$iso[country.vars$poor == 1])
rich.ma30.cce <- lr_het_coef(het.coeff.cce,country.vars$iso[country.vars$aeg20em == 1])

mytab <- rbind(c("\\hat{\\theta}",formatC(c(all.ma30.nocce$lr_coeff,all.ma30.cce$lr_coeff),digits = 3,format = "f")),
               c("",paste0("(",formatC(c(all.ma30.nocce$std_err,all.ma30.cce$std_err),digits = 3,format = "f"),")")),
               c("\\hat{\\theta}",formatC(c(cold.ma30.nocce$lr_coeff,cold.ma30.cce$lr_coeff),digits = 3,format = "f")),
               c("",paste0("(",formatC(c(cold.ma30.nocce$std_err,cold.ma30.cce$std_err),digits = 3,format = "f"),")")),
               c("\\hat{\\theta}",formatC(c(hot.ma30.nocce$lr_coeff,hot.ma30.cce$lr_coeff),digits = 3,format = "f")),
               c("",paste0("(",formatC(c(hot.ma30.nocce$std_err,hot.ma30.cce$std_err),digits = 3,format = "f"),")")),
               c("\\hat{\\theta}",formatC(c(poor.ma30.nocce$lr_coeff,poor.ma30.cce$lr_coeff),digits = 3,format = "f")),
               c("",paste0("(",formatC(c(poor.ma30.nocce$std_err,poor.ma30.cce$std_err),digits = 3,format = "f"),")")),
               c("\\hat{\\theta}",formatC(c(rich.ma30.nocce$lr_coeff,rich.ma30.cce$lr_coeff),digits = 3,format = "f")),
               c("",paste0("(",formatC(c(rich.ma30.nocce$std_err,rich.ma30.cce$std_err),digits = 3,format = "f"),")")))
               
colnames(mytab) <- c("~","No CCE","CCE")

knitr::kable(mytab,digits=3,align = c("l", rep("c", 2)), escape = FALSE, format = "html") |>
  kableExtra::group_rows("All countries", 1, 2, hline_after = TRUE) |>
  kableExtra::group_rows("Cold (Average temperature below 33rd percentile)", 3, 4, hline_after = TRUE) |>
  kableExtra::group_rows("Temperate or hot (Average temperature above 33rd percentile)", 5, 6, hline_after = TRUE) |>
  kableExtra::group_rows("Poor (Low income developing countries)", 7, 8, hline_after = TRUE) |>
  kableExtra::group_rows("Rich (Advanced Economies and G20 Emerging Markets)", 9, 10 , hline_after = TRUE) |>
     kableExtra::kable_styling(font_size = 10) 
```
## References